Malacology: A Programmable Storage System Built on Ceph

### 2016-05-09 paper/paper.tex
modified:   paper.tex 	Polished section 2

\begin{figure}[tb]
\centering
\includegraphics{figures/overview.png}
\caption{A Ceph installation consists of a cluster running mostly
storage daemons (OSDs) to ensure data durability, a few monitor daemons
(MONs) to reach consensus of the cluster state and to ensure its
consistent versioning, and a few metadata service daemons (MDSs) for
scalable metadata service. Ceph's APIs include object, block, file
abstractions. Malacology enables the programmability of these
abstractions and services (indicated by bold arrows). To illustrate Malacology we implement two
services on Ceph, Zlog and Mantle, by programming Ceph's durability,
consistent versioning, consensus, and metadata subsystems.
\label{fig:overview}}
\end{figure}

@@

@@

So called ``bolt-on'' services are third-party solutions integrated into a
system to provide functionality and improve performance for common use-cases
(e.g. a metadata service). The introduction of additional services may provide
simplicity and performance enhancements, but ``bolt-on'' services often
duplicate functionality, unnecessarily increasing the likelihood of bugs or
introducing unpredictable performance problems. For example, we are developing
a high-performance distributed shared-log on Ceph called ZLog, based on the
CORFU protocol~\cite{corfu}. Serialization at high-performance is achieved using a
``sequencer'' service that maintains a volatile, in-memory counter used to assign a log
position to each client that appends to the log. One option for building
a distributed log is to use PAXOS, or an out-of-the-box solution such as
Zookeeper. However, these systems adhere to the pre-coneived notions of
storage, namely that data is safe from data loss, which is exactly the feature
that causes traditional approaches to run at significantly reduced speed.

@@

For example, SciHadoop~\cite{buck:hpc2011-scihadoop,buck:sc2013-scidr} changes both the Hadoop
application and the Hadoop framework itself to leverage the structure of
scientific data (3D array-based data, more specifically) to increase
performance, locality, and the number of early results. This is not a
bad proposition, but creates a coupling that is highly tied to the
underlying physical properties of the system, making it difficult to
adapt to future changes at the storage system level.

@@

When these two approaches fail to meet the needs of the application,
developers turn their attention to the storage system itself.
Traditionally, storage system behaviour can be altered using two
techniques: tuning the system or introducing changes to the system. The
difficulty of both of these approaches has given rise to a third
technique called active storage.

@@

Tuning large and complicated systems is difficult because of the number
of parameters (e.g., Hadoop 2.7.1 has 210 tunables) and because the
tunables are difficult to understand or quantify (e.g., Ceph tunables~\cite{sevilla:sc15-mantle}). To succesfully tune the system, a
developer must have domain and system specific knowledge. Without this
intimate knowledge, the tuning turns into somewhat of a black art, where
the only way to figure out the best settings is trial and error.

@@

Alternatively, some systems have formal support for introducing additional
functionality, often in the form of a plugin infrastructure. Ceph provides the
ability to create domain-specific logical object interfaces by composing
existing interfaces. These interfaces are written in C++ and are statically
loaded into the system. An example usage may be to atomically combine a
``write'' with an index update to create a content-addressable object
interface.

@@

Figure 1 shows the throughput
(\emph{y} axis) over time (\emph{x} axis) of two OSD implementations for
storing metadata: in the header of the byte stream (data) vs.~in the
object extended attributes (XATTR). The speed for appending data without
any metadata operations (data raw) is also shown as a baseline for
comparison. For append-heavy workloads storing metadata as a header in
the data performs about 1.5x better than storing metadata as an extended
attribute.

\begin{figure}[htbp]
\centering
\includegraphics{figures/cls_xattr-vs-data.png}
\caption{{[}\href{https://github.com/michaelsevilla/malacology-popper/blob/master/experiments/figure1/visualize.ipynb}{source}{]}
When appending data to objects, the object class that stores metadata in
a header in the data byte stream (data) performs 1.5x better than the
object class that stores metdata in the extended attributes of the
object (XATTR); it is almost as fast as appending data without updating
metadata (data raw).}
\end{figure}

The object interface facility provided by Ceph is more than an
a proof-of-concept active storage implementation.  Figure
\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
domain-specific interfaces in the Ceph community since 2010.  What is most
remarkable is that this trend contradicts the notion that API changes are a
burden for users.
Rather it appears that a gap in existing interfaces are
being addressed through programmability. In fact, Table~\ref{table:objclasses}
categorizes existing interfaces and we clearly see a trend towards reusable
services.

\begin{figure}[ht]
\centering
\includegraphics{figures/obj-int-dev-growth.png}
\caption{Since 2010, the growth in the number of co-designed object
storage interfaces in Ceph has been accelerating. This plot is the
number of object classes (a group of interfaces), and the total number
of methods (the actual API end-points).\label{fig:obj-int-dev-growth}}
\end{figure}

@@

The popularity of the active storage component of Ceph hint at three
trends in the Ceph community: (1) increasingly, the default
algorithms/tunables of the storage system are insufficient for the
application's performance goals, (2) programmers are becoming more aware
of their application's behavior, and (3) programmers know how to manage
resources to improve performance. Programmers gravitate towards object
interfaces because it gives them ability to tell the storage system
about their application: if it is CPU or IO bound, if it has locality,
if its size has the potential to overload a single proxy node, etc. The
programmers know what the problem is and how to solve it, but until
object interfaces, had no way to tell the storage system how to handle
their data.

These implementations use the active storage interfaces in Ceph. We
refer to this as \emph{storage programmability} which is a method by
which an application communicates its requirements to the storage system
in a way that allows the application to realize a new behavior without
sacrificing the correctness of the underlying system. This evaluation
and resulting implementation would have been a burden without object
classes.

@@

\section{Services Built on Malacology}\label{services-built-on-malacology}
\label{services}

\begin{table}
\centering
\begin{tabular}{  l | l | l    }
\textbf{Service} & \textbf{Ceph Component} & \textbf{Used for...}  \\ \hline
Mantle  & RADOS    & policy durability \\
        & MON maps & policy versioning \\ \hline
ZLog    & Logical Storage & CORFU protocol  \\ 
        & MON maps & interface install \\ \hline
Seq.    & MDS file types & log management \\ 
        & MDS fault toler. & seq. fault toler. \\
\end{tabular}
\caption{To implement the Mantle and ZLog services, we leverage Ceph components. Malacology uses these subsystems because they provide powerful abstractions.}
\label{table:implementation}
\end{table}

>>>>

@@

\begin{figure}[htbp]
\centering
\includegraphics{figures/overview.png}
\caption{A Ceph installation consists of a cluster running mostly
storage daemons (OSDs) to ensure data durability, a few monitor daemons
(MONs) to reach consensus of the cluster state and to ensure its
consistent versioning, and a few metadata service daemons (MDSs) for
scalable metadata service. Ceph's APIs include object, block, file
abstractions. Malacology enables the programmability of these
abstractions and services. To illustrate Malacology we implement two
services on Ceph, Zlog and Mantle, by programming Ceph's durability,
consistent versioning, consensus, and metadata subsystems.
\label{fig:overview}}
\end{figure}

@@

So called ``bolt-on'' services are third-party solutions integrated into a
system to provide functionality and improve performance for common use-cases
(e.g. a metadata service). The introduction of additional services may provide
simplicity and performance enhancements, but ``bolt-on'' services often
duplicate functionality, unnecessarily increasing the likelihood of bugs or
introducing unpredictable performance problems. For example, we are developing
a high-performance distributed shared-log on Ceph called ZLog, based on the
CORFU protocol~\cite{corfu}. Without going into greater detail, the CORFU protocol by separating the concern of correctness from the concern of performance enables serialization using a \emph{volatile}
``sequencer'' service. This high-performance in-memory counter is used to assign a log
position to each client that appends to the log. The sequencer is purely a performance optimization and does not have a role in correctness (if the sequencer fails, the tail of the log can always be retrieved by other means). Thus, it is the volatiity of the sequence that enables high performance. A ``bolt-on'' option for building
a distributed log is to use PAXOS, or an out-of-the-box solution such as
Zookeeper. However, these systems adhere to the pre-coneived notions of
storage, namely that data is safe from data loss, which is exactly the feature
that causes traditional approaches to run at significantly reduced speed.

@@

For example, SciHadoop~\cite{buck:hpc2011-scihadoop,buck:sc2013-scidr} changes both the Hadoop
application and the Hadoop framework itself to leverage the structure of
scientific data (3D array-based data, more specifically) to increase
performance, locality, and the number of early results. %This is not a
%bad proposition, but 
Changing applications creates a coupling that is highly tied to the
underlying physical properties of the system, making it difficult to
adapt to future changes at the storage system level.

@@

When both the bolt-on approach and changing applications fail to meet the needs of the application,
developers turn their attention to the storage system itself.
Traditionally, storage system behaviour can be altered using two
techniques: tuning the system or introducing changes to the system. The
difficulty of both of these approaches has given rise to a third
technique called active storage.

@@

Tuning large and complicated systems is difficult because of the number
of parameters (e.g., Hadoop 2.7.1 has 210 tunables) and because the
tunables are difficult to understand or quantify (e.g., Ceph tunables~\cite{sevilla:sc15-mantle}). To succesfully tune the system, a
developer must have domain and system-specific knowledge. Without this
intimate knowledge, the tuning turns into somewhat of a black art, where
the only way to figure out the best settings is trial and error.

@@

Alternatively, some systems have formal support for introducing additional
functionality, often in the form of a plugin infrastructure. Ceph provides the
ability to create domain-specific logical object interfaces by composing
existing interfaces. These interfaces are written in C++ and are statically
loaded into the system. %An example usage may be to atomically combine a
%``write'' with an index update to create a content-addressable object
%interface.

@@

For example, Figure~\ref{fig:datavsxattr} compares the throughput
(\emph{y} axis) over time (\emph{x} axis) of two OSD implementations of the same storage object interface for
storing metadata: either in the header of the byte stream (data) or in the
object extended attributes (XATTR). The speed for appending data without
any metadata operations (data raw) is also shown as a baseline for
comparison. For append-heavy workloads storing metadata as a header in
the data performs about 1.5x better than storing metadata as an extended
attribute.

\begin{figure}[htbp]
\centering
\includegraphics{figures/cls_xattr-vs-data.png}
\caption{{[}\href{https://github.com/michaelsevilla/malacology-popper/blob/master/experiments/figure1/visualize.ipynb}{source}{]}
When appending data to objects, the object class that stores metadata in
a header in the data byte stream (data) performs 1.5x better than the
object class that stores metdata in the extended attributes of the
object (XATTR); it is almost as fast as appending data without updating
metadata (data raw).}
\label{fig:datavsxattr}
\end{figure}

The object interface facility provided by Ceph is more than an
a proof-of-concept active storage implementation.  Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
domain-specific interfaces in the Ceph community since 2010.  What is most
remarkable is that this trend contradicts the notion that API changes are a
burden for users.
Rather it appears that a gap in existing interfaces are
being addressed through programmability. In fact, Table~\ref{table:objclasses}
categorizes existing interfaces and we clearly see a trend towards reusable
services.

\begin{figure}[ht]
\centering
\includegraphics{figures/obj-int-dev-growth.png}
\caption{Since 2010, the growth in the number of co-designed object
storage interfaces in Ceph has been accelerating. This plot is the
number of object classes (a group of interfaces), and the total number
of methods (the actual API end-points).}
\label{fig:obj-int-dev-growth}
\end{figure}

@@

The popularity of the active storage component of Ceph hint at three
trends in the Ceph community: (1) increasingly, the default
algorithms/tunables of the storage system are insufficient for the
application's performance goals, (2) programmers are becoming more aware
of their application's behavior, and (3) programmers know how to manage
resources to improve performance. Programmers gravitate towards object
interfaces because it gives them the ability to tell the storage system
about their application: if it is CPU or IO bound, if it has locality,
if its size has the potential to overload a single proxy node, etc. The
programmers know what the problem is and how to solve it, but until
the ability to modify object interfaces, they had no way to express to the storage system how to handle
their data.

\if{false}
These implementations use the active storage interfaces in Ceph. We
refer to this as \emph{storage programmability} which is a method by
which an application communicates its requirements to the storage system
in a way that allows the application to realize a new behavior without
sacrificing the correctness of the underlying system. This evaluation
and resulting implementation would have been a burden without object
classes.
\fi

@@

\section{Services Built on
Malacology}\label{services-built-on-malacology}

\label{services}

\begin{table}
\begin{tabular}{ | l | l | l | l | l |}
\hline
Cap Owner & Failure & ZLog Recovery \\ \hline
mds & mds & mds \\ \hline
client & mds & xfer client cap \\ \hline
mds & client & n/a \\ \hline
client & client & mds \\
\hline
\end{tabular}
\caption{Failure and recovery scenarios.}
\label{t:mds-zlog-fail}
\end{table}
